Brother, Can You Spare a Manufacturing Job? How Voters React to Deindustrialization

### 2021-12-13 Final Draft/Final Draft.Rmd
updates

This paper analyzes the relationship between deindustrialization, race, and the election of Donald Trump in 2016. Previous analysis has  focused on racial animus, trade, or the conjunction of the two on  political polarization and election outcomes (\cite{Autor20}; \cite{Che16}; \cite{BR21}). We extend the analysis of \cite{Baccini21}, one of the first papers to examine the political effects of deindustrialization that while considering race and localization. Deindustrialization - the transition of the U.S. economy away from manufacturing - has led to widespread job loss and created adverse knock-on effects for a wide swath of Americans. Our measure of deindustrialization incorporates job losses due to trade with China and others (see \cite{Acemoglu16}); it also includes job losses due to automation, demand changes, and other factors. The experience of deindustrialization has not been uniform; rather, it has been concentrated in certain areas of the country that were previously economically reliant on manufacturing, as demonstrated in Figure \ref{ManuMap}. The localized nature of this analysis highlights the uneven and concentrated harms of deindustrialization.

@@

We want to measure the causal relationship between deindustrialization and the change in vote share. As a measure of deindustrialization, we use the loss of manufacturing jobs as a share of the total beginning-of-period employment in each county. Suppose at the beginning of our period there were $2000$ manufacturing workers in a county, and $8000$ non-manufacturing workers. At the end there are only $1500$ manufacturing workers. Using our measure of net change in manufacturing employment, this is a loss of $500/10000 = 5\%.$

In this choice we diverge from \cite{Baccini21}, which uses *gross* manufacturing job losses. For example, if a county lost $450$ manufacturing jobs from 2012-2015 and also gained $400$ manufacturing jobs over the same period, this would be $450$ *gross* job losses, but only $50$ *net* job losses. The *gross* measure captures several dynamics unrelated to deindustrialization - for example, seasonal unemployment in a food-manufacturing region, or workers moving between jobs. Thus we believe the *net* job losses over a period are the more accurate measure. Additionally, there is precedent in the literature to use *net* employment changes, as \cite{Autor21} have done. Nevertheless, it can be argued that the non-seasonal or non-structural components of gross job losses are a useful covariate. For instance, basic notions of prospect theory suggest that the loss of a job and subsequent regaining of another identical job may still create substantial discontent. Constructing a better measure of gross job losses is an area for further methodological and substantive research.

@@

We conduct two related tests of our hypothesis. Our first test involves an instrumental-variable approach. A regression of change in Democratic vote share on manufacturing job losses risks endogeneity, in case counties with manufacturing job losses were otherwise predisposed to turn towards Trump (a singular candidate, after all). To mitigate these risks, \cite{Baccini21} use a Bartik instrument, which we adapt (see \cite{Bartik91}). This instrument essentially uses the cross-county distribution of manufacturing employment as a source of exogenous variation.

@@

Our second approach involves attempting to balance the covariates of the deindustrialization treatment. We use the same covariates as are used in controls above, as well as manufacturing as share of the population and total population. We apply the method of Covariate Balancing Propensity Scores, as developed by \cite{Imai14}; this method generates weights to reduce the correlation between covariates and treatment, while addressing several of the issues with propensity score weights. We find significant imbalance in the data - in other words, it is relatively easy to predict the 'treatment' of job loss given the covariates (see Appendix for balance plots). Given covariate-balancing weights, we then conduct a single-stage weighted regression of change in Democratic vote share on manufacturing job loss, using the same controls as above.

>>>>

@@

We want to measure the causal relationship between deindustrialization and the change in vote share. As a measure of deindustrialization, we use the loss of manufacturing jobs as a share of the total beginning-of-period employment in each county. Suppose at the beginning of our period there were $2000$ manufacturing workers in a county, and $8000$ non-manufacturing workers. At the end there are only $1500$ manufacturing workers. Using our measure of net manufacturing job losses, this is a loss of $500/10000 = 5\%.$

In this choice we diverge from \cite{Baccini21}, which uses *gross* manufacturing job losses. For example, if a county lost $450$ manufacturing jobs from 2012-2015 and also gained $400$ manufacturing jobs over the same period, this would be $450$ *gross* job losses, but only $50$ *net* job losses. The *gross* measure captures several dynamics unrelated to deindustrialization - for example, seasonal unemployment in a food-manufacturing region, or workers moving between jobs. Thus we believe the *net* job losses over a period are the more accurate measure of deindustrialization. Additionally, there is precedent in the literature to use *net* employment changes, as \cite{Autor21} have done. Nevertheless, it can be argued that the non-seasonal or non-structural components of gross job losses are a useful covariate. For instance, basic notions of prospect theory suggest that the loss of a job and subsequent regaining of another identical job may still create substantial discontent. Constructing a better measure of gross job losses is an area for further methodological and substantive research.

@@

We conduct two related tests of our hypothesis. Our first test involves an instrumental-variable approach. A regression of change in Democratic vote share on manufacturing job losses risks endogeneity, in case counties with manufacturing job losses were otherwise predisposed to turn towards Trump. To mitigate these risks, \cite{Baccini21} use a Bartik instrument, which we adapt (see \cite{Bartik91}). This instrument essentially uses the cross-county distribution of manufacturing employment as a source of exogenous variation.

@@

A potential weakness of our first approach is that manufacturing employment may be correlated with swinging towards Trump for reasons other than manufacturing job losses. To account for this potential ommitted variable bias, we use a second approach that attempts to balance the covariates of the deindustrialization treatment. We use the same covariates as are used in controls above, as well as manufacturing as share of the population and total population. We apply the method of Covariate Balancing Propensity Scores, as developed by \cite{Imai14}; this method generates weights to reduce the correlation between covariates and treatment, while addressing several of the issues with propensity score weights. We find significant imbalance in the data - in other words, it is relatively easy to predict the 'treatment' of job loss given the covariates (see Appendix for balance plots). Given covariate-balancing weights, we then conduct a single-stage weighted regression of change in Democratic vote share on manufacturing job loss, using the same controls as above. Essentially, this approach can be conceptualized as a form of matching that reduces the potential ommitted variable bias in our first approach, but may introduce selection bias (as some counties will be given low weights).

### 2021-12-13 Final Draft/Final-Draft.tex
updates

We want to measure the causal relationship between deindustrialization
and the change in vote share. As a measure of deindustrialization, we
use the loss of manufacturing jobs as a share of the total
beginning-of-period employment in each county. Suppose at the beginning
of our period there were \(2000\) manufacturing workers in a county, and
\(8000\) non-manufacturing workers. At the end there are only \(1500\)
manufacturing workers. Using our measure of net change in manufacturing
employment, this is a loss of \(500/10000 = 5\%.\)

In this choice we diverge from \cite{Baccini21}, which uses \emph{gross}
manufacturing job losses. For example, if a county lost \(450\)
manufacturing jobs from 2012-2015 and also gained \(400\) manufacturing
jobs over the same period, this would be \(450\) \emph{gross} job
losses, but only \(50\) \emph{net} job losses. The \emph{gross} measure
captures several dynamics unrelated to deindustrialization - for
example, seasonal unemployment in a food-manufacturing region, or
workers moving between jobs. Thus we believe the \emph{net} job losses
over a period are the more accurate measure. Additionally, there is
precedent in the literature to use \emph{net} employment changes, as
\cite{Autor21} have done. Nevertheless, it can be argued that the
non-seasonal or non-structural components of gross job losses are a
useful covariate. For instance, basic notions of prospect theory suggest
that the loss of a job and subsequent regaining of another identical job
may still create substantial discontent. Constructing a better measure
of gross job losses is an area for further methodological and
substantive research.

@@

Our second approach involves attempting to balance the covariates of the
deindustrialization treatment. We use the same covariates as are used in
controls above, as well as manufacturing as share of the population and
total population. We apply the method of Covariate Balancing Propensity
Scores, as developed by \cite{Imai14}; this method generates weights to
reduce the correlation between covariates and treatment, while
addressing several of the issues with propensity score weights. We find
significant imbalance in the data - in other words, it is relatively
easy to predict the `treatment' of job loss given the covariates (see
Appendix for balance plots). Given covariate-balancing weights, we then
conduct a single-stage weighted regression of change in Democratic vote
share on manufacturing job loss, using the same controls as above.

>>>>

We want to measure the causal relationship between deindustrialization
and the change in vote share. As a measure of deindustrialization, we
use the loss of manufacturing jobs as a share of the total
beginning-of-period employment in each county. Suppose at the beginning
of our period there were \(2000\) manufacturing workers in a county, and
\(8000\) non-manufacturing workers. At the end there are only \(1500\)
manufacturing workers. Using our measure of net manufacturing job
losses, this is a loss of \(500/10000 = 5\%.\)

In this choice we diverge from \cite{Baccini21}, which uses \emph{gross}
manufacturing job losses. For example, if a county lost \(450\)
manufacturing jobs from 2012-2015 and also gained \(400\) manufacturing
jobs over the same period, this would be \(450\) \emph{gross} job
losses, but only \(50\) \emph{net} job losses. The \emph{gross} measure
captures several dynamics unrelated to deindustrialization - for
example, seasonal unemployment in a food-manufacturing region, or
workers moving between jobs. Thus we believe the \emph{net} job losses
over a period are the more accurate measure of deindustrialization.
Additionally, there is precedent in the literature to use \emph{net}
employment changes, as \cite{Autor21} have done. Nevertheless, it can be
argued that the non-seasonal or non-structural components of gross job
losses are a useful covariate. For instance, basic notions of prospect
theory suggest that the loss of a job and subsequent regaining of
another identical job may still create substantial discontent.
Constructing a better measure of gross job losses is an area for further
methodological and substantive research.

@@

A potential weakness of our first approach is that manufacturing
employment in a county may be correlated with swinging towards Trump for
reasons other than manufacturing job losses. To account for this
potential ommitted variable bias, we use a second approach that attempts
to balance the covariates of the deindustrialization treatment. We use
the same covariates as are used in controls above, as well as
manufacturing as share of the population and total population. We apply
the method of Covariate Balancing Propensity Scores, as developed by
\cite{Imai14}; this method generates weights to reduce the correlation
between covariates and treatment, while addressing several of the issues
with propensity score weights. We find significant imbalance in the data
- in other words, it is relatively easy to predict the `treatment' of
job loss given the covariates (see Appendix for balance plots). Given
covariate-balancing weights, we then conduct a single-stage weighted
regression of change in Democratic vote share on manufacturing job loss,
using the same controls as above. Essentially, this approach can be
conceptualized as a form of matching that reduces the potential ommitted
variable bias in our first approach, but may introduce selection bias
(as some counties will be given low weights).
